\section{Overall Description}

\subsection{Product Perspective}

The Clinic Management System is a standalone software system designed to replace traditional paper-based processes with a fully automated digital environment. While it operates independently, the system is designed to interact within a modern technical ecosystem:

\begin{itemize}
	\item \textbf{User Interface:} A responsive web application accessible via modern web browsers.
	\item \textbf{Hardware Integration:} The system supports interaction with standard office printers for generating invoices, prescriptions, and medical reports.
	\item \textbf{External Systems (Future Integration):} The database architecture is designed to support future scalability, allowing integration with third-party services such as SMS gateways for appointment reminders and electronic payment portals.
\end{itemize}

\subsection{Product Functions}

The system provides a set of core services organized into the following functional modules:

\begin{itemize}
	\item \textbf{Identity and Access Management (IAM):} Controlling user authentication and Role-Based Access Control (RBAC).
	\item \textbf{Appointment Scheduling:} Booking, rescheduling, and canceling appointments with automated conflict detection.
	\item \textbf{Medical Records Management:} Digitally archiving patient history, diagnoses, and prescriptions.
	\item \textbf{Billing and Finance:} Automated invoice generation and tracking of payments.
	\item \textbf{Audit and Compliance:} Maintaining an active log of all system activities for transparency and security.
\end{itemize}

\subsection{User Classes and Characteristics}

The system identifies four primary user classes:

\begin{itemize}
	\item \textbf{System Administrator (Admin):} Full system privileges including user management and audit logs.
	\item \textbf{Receptionist:} Manages patients, appointments, and billing operations.
	\item \textbf{Doctor:} Handles medical records, diagnoses, and schedules.
	\item \textbf{Patient:} Can request appointments and view personal medical data.
\end{itemize}

\subsection{Operating Environment}

The system operates within a distributed web architecture:

\begin{itemize}
	\item \textbf{Platform:} Web-based application deployed on a centralized server.
	\item \textbf{Client Side:} Accessible via PCs, laptops, or tablets using modern browsers.
	\item \textbf{Operating System:} Compatible with Linux and Windows Server environments.
	\item \textbf{Database:} MySQL relational database management system.
\end{itemize}

\subsection{Design and Implementation Constraints}

The system must respect the following constraints:

\begin{itemize}
	\item \textbf{Technical Stack:} PHP (backend), Vue.js (frontend), HTML5/CSS3/JavaScript, MySQL database.
	\item \textbf{Security and Privacy:} Strict access control for medical data based on user roles.
	\item \textbf{Availability:} Minimum 99\% uptime during clinic working hours.
\end{itemize}

\subsection{User Documentation}

The system includes:

\begin{itemize}
	\item User Manual (Arabic \& English PDF).
	\item Administrator Guide (system configuration and backup procedures).
	\item In-app help tooltips and guidance messages.
\end{itemize}

\subsection{Assumptions and Dependencies}

\textbf{Assumptions:}
The clinic provides hardware devices and stable internet connectivity. Printers support standard web printing protocols.

\textbf{Dependencies:}
The system depends on MySQL database availability and web hosting infrastructure.