\section{External Interface Requirements}

\subsection{User Interfaces}

The system features a modern Graphical User Interface (GUI) developed with Vue.js to provide a Single Page Application (SPA) experience.

\begin{itemize}
	\item \textbf{Dashboards:} Specialized views for Doctors, Receptionists, and Patients to minimize cognitive load.
	\item \textbf{UX/UI Standards:} Consistent typography and color palettes designed for long-term daily use.
	\item \textbf{Validation:} Real-time client-side validation and confirmation dialogs for critical actions (e.g., deleting records).
\end{itemize}

\subsection{Hardware Interfaces}

The system supports direct interfacing with the following hardware:

\begin{itemize}
	\item \textbf{Printers:} For generating physical copies of medical bills, prescriptions, and laboratory reports via the browser's print subsystem.
\end{itemize}

\subsection{Software Interfaces}

\begin{itemize}
	\item \textbf{DBMS Interface:} Utilizing PHP PDO to manage secure communication between the application logic and the MySQL/Oracle database.
	\item \textbf{APIs:} Implementing RESTful APIs to serve JSON data from the PHP backend to the Vue.js frontend.
	\item \textbf{Payment Integration:} Future hooks for external Payment Gateways to process transactions within the Payments table.
\end{itemize}

\subsection{Communications Interfaces}

\begin{itemize}
	\item \textbf{Protocols:} Mandatory use of HTTPS for end-to-end data encryption.
	\item \textbf{Network Support:} Compatible with both Local Area Networks (LAN) and Wide Area Networks (WAN).
	\item \textbf{Data Interchange:} All data exchange between the server and client will be formatted in JSON.
\end{itemize}