\section{Other Non-functional Requirements}

\subsection{Performance}
\begin{itemize}
	\item \textbf{Response Time:} The system shall respond to user queries (e.g., patient search, appointment check) within a maximum of 2 seconds.
	\item \textbf{Concurrency:} The system must support at least 50 concurrent users without degradation in performance, especially during peak clinic hours.
\end{itemize}

\subsection{Security}
\begin{itemize}
	\item \textbf{Authentication:} Access is restricted to authorized users via encrypted credentials with Role-Based Access Control (RBAC).
	\item \textbf{Encryption:} Sensitive data, including passwords and payment information, must be encrypted using strong protocols such as BCrypt or AES.
	\item \textbf{Authorization:} The system follows the Principle of Least Privilege.
	\item \textbf{Audit Trail:} The \texttt{Active\_log} table shall record every operation (Create, Update, Delete) with timestamp and user ID.
\end{itemize}

\subsection{Usability}
\begin{itemize}
	\item \textbf{Intuitive UI:} Designed to ease transition from paper-based processes with minimal training.
	\item \textbf{Multi-language Support:} English primary, Arabic supported, future French expansion.
\end{itemize}

\subsection{Reliability \& Availability}
\begin{itemize}
	\item \textbf{Uptime:} 99.9\% availability during clinic operational hours.
	\item \textbf{Data Recovery:} Mechanisms to prevent data loss during failures.
\end{itemize}

\subsection{Maintainability \& Architecture}
\begin{itemize}
	\item \textbf{MVC Architecture:}
	\begin{itemize}
		\item Front-End: HTML, CSS, Vue.js
		\item Logic Layer: PHP Controllers
		\item Data Access Layer: OOP PHP abstraction for DBMS flexibility
	\end{itemize}
	\item \textbf{Coding Standards:} CamelCase, documentation, SOLID principles.
\end{itemize}

\subsection{Compatibility}
\begin{itemize}
	\item \textbf{Browser Support:} Chrome, Firefox, Edge, Safari.
	\item \textbf{Responsive Design:} Desktop, Tablet, Smartphone support.
\end{itemize}

\subsection{Health Data Governance}

\subsubsection{Data Classification}
PII, PHI, Financial Data, System Logs.

\subsubsection{Access Control}
Role-based restrictions and logging of PHI access.

\subsubsection{Data Encryption}
Encryption at rest and HTTPS/TLS in transit.

\subsubsection{Data Retention Policy}
Medical records (10 years), audit logs (5 years).

\subsubsection{Patient Privacy \& Consent}
Patient consent, data access, and correction rights.

\subsubsection{Backup \& Disaster Recovery}
Daily backups, 2-hour restore window, off-site encrypted storage.

\subsubsection{Audit \& Monitoring}
All PHI operations logged with User ID, Timestamp, Action.

\subsection{Operational Policies}

\subsubsection{Backup Policy}
Daily encrypted backups, 30-day rolling retention, 2-hour restore.

\subsubsection{Session Management Policy}
15-minute timeout, restricted concurrent logins, secure tokens.

\subsubsection{Password \& Authentication Policy}
8+ chars, bcrypt hashing, 90-day rotation, lockout after 5 failures.

\subsubsection{System Monitoring Policy}
Critical operation logs, admin review, suspicious login alerts.